
---
title: "Daily arXiv Articles in the areas of Data Science"
description: |
  There are 60 articles published today.
author:
  - name: Bryan Whiting
    url: https://www.bryanwhiting.com
date: 2019-12-18
output:
  distill::distill_article:
    self_contained: false
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
install.load::install_load('DT', 'dplyr', 'tidyverse', 'readr', 'stringr', 'kableExtra')


# Tags I care about
tags <- c('stat.AP', 'stat.co', 'stat.ML', 'cs.LG', 'q-fin.ST', 'q-fin.EC', 'econ-EM')
LINK_COLOR = "#d9230f"

create_subjects <- function(df){
  df %>% 
    mutate(subject = str_extract(primary_tag, '^[a-z\\-]*'),
         subject = case_when(
           subject == 'cs' ~ 'Computer Science',
           subject == 'physics' ~ 'Physics',
           subject == 'econ' ~ 'Economics',
           subject == 'eess' ~ 'Elec. Eng. and Systems Science',
           subject == 'math' ~ 'Mathematics',
           subject == 'q-bio' ~ 'Quantitative Biology',
           subject == 'q-fin' ~ 'Quantitative Finance',
           subject == 'stat' ~ 'Statistics',
           subject == 'quant-ph' ~ 'Quantum Physics',
           subject == 'cond-mat' ~ 'Condensed Matter',
           TRUE ~ 'Other'
         )) %>%
    return()
}

df = read_csv('arxiv.csv') 
df <- df %>% 
  create_subjects()


prep_abstracts <- function(df){
  #' This function prepares a dataframe with html
  #' 
  df %>%
  # TODO: Sort by number of authors. FUTURE: sort by pagerank of authors
  arrange(desc(n_authors)) %>%
  mutate(
    primary_category = paste0(primary_category, ' (',primary_tag,')'),
    # Find the index for the nth period
    idx_period = str_locate(summary, '^([^.]+\\.){2}')[,2],
    title = paste0('<b><a href=',url_href ,' style="color: #d9230f">', title, '</a></b>',
                   '<br><em>', categories,  '</em>', 
                   '. ', n_authors, ' authors. ',
                   '<a href=', url_pdf, ' style="color: #d9230f;
    text-decoration: none;">pdf</a>', 
    # Create the abstract summary
                   '<br><details><summary>', str_sub(summary, 1, idx_period), ' ...</summary><br>',
                   str_sub(summary, idx_period + 1), '</details>')) %>%
  select(primary_category, title) 
}
#df %>% filter(subject == 'Statistics') %>% prep_abstracts()

create_html_table <- function(df){
  #' This function groups rows and prepares a pretty able of the data from prep_abstracts
  
  # Identify the break points in the Category
  # Grouping rows via labeling: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#group_rows_via_labeling
  # breakpoints for the pack_rows
  breaks <- df %>% 
    group_by(primary_category) %>% 
    count(sort=T) %>% 
    ungroup() %>%
    mutate(end = cumsum(n),
           start = end - n + 1) %>%
    select(primary_category, start, end, n) 
  
  tab <- df %>%
    select(-primary_category) %>%
    kable(escape=F, format='html', col.names=NULL) %>%
    kable_styling(bootstrap_options = c('striped', 'condensed')) 
  
  # This could be simplified with index = c(), but I'm lazy
  for (i in 1:nrow(breaks)){
    tab <- tab %>%
      pack_rows(breaks$primary_category[i], 
                breaks$start[i],
                breaks$end[i],
                label_row_css = "background-color: #666; color: #fff;")
  } 
  return(tab)
}
# df %>% 
#   filter(subject == 'Statistics') %>% 
#   prep_abstracts() %>% 
#   create_html_table()
    
```


## Breakdown of arXiv Publication Counts
Yesterday's counts of submitted papers on www.arxiv.org grouped by primary subject. Click the links in the table to be re-directed to the abstracts below. The links under `Subject` will redirect you to abstracts with the primary subject (there can only be one primary subject on arXiv). The links under `Category` will redirect you to all publications yesterday with a given tag (primary or secondary).

```{r, fig.cap='Breakdown of arXiv categories by Primary Tag'}


df_primary <- df %>% 
  group_by(primary_category, primary_tag) %>% 
  count(sort=TRUE)
# filter all the other tags
df_other <- df %>% 
  # remove the primary tag
  mutate(tags = str_replace(tags, primary_tag, '')) %>% 
  # Create a single vector
  summarize(paste(tags, collapse=', ')) %>% 
  pull() %>% 
  str_split(', ') %>%  # this returns a list
  `[[`(1) %>% # extract first element in list
  data.frame(category = .) %>% # convert back to dataframe
  filter(category != "") %>% # remove empty tags
  group_by(category) %>% 
  count(sort=T) %>%
  ungroup() %>%
  mutate(category = as.character(category)) %>%
  rename(n_secondary=n) 

df_summary <- df_primary %>% 
  create_subjects() %>%
  rename(category=primary_tag) %>% 
  left_join(df_other, by='category') %>%
  # Arrange the tables by group max
  group_by(subject) %>%
  mutate(n_tot = sum(n)) %>%
  arrange(desc(n_tot), subject, desc(n), desc(n_secondary)) %>%
  ungroup() %>%
  select(subject, n_tot, everything())

# TODO: Find a way to improve the description by printing this out and having it be in the title
# Save out data for tweets

# df_summary %>% 
#   mutate(subject = paste0(subject, ' (', n_tot, ')')) %>%
#   select(subject) %>%
#   unique() %>% pull() %>% paste(collapse = ', ')
  

# collapse the info across rows
df_summ_formatted <- df_summary %>%
  # the text_spec provides an anchor: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#links
  mutate(
    compact_subject = paste0(subject, ' (', n_tot, ')'),
    subj_link = paste0('#', str_replace(tolower(subject), fixed(' '), '-')),
    subject = text_spec(compact_subject, link=subj_link, color=LINK_COLOR, bold=T),
    tag = category,
    # format the category 
    compact_category = paste0(primary_category, ' (', category, ')'),
    cat_link = paste0('#', tolower(compact_category)),
    cat_link = str_replace_all(cat_link, fixed(' '), '-'),
    cat_link = str_replace_all(cat_link, fixed('('), '-'),
    cat_link = str_replace_all(cat_link, fixed(')'), '-'),
    cat_link = str_replace_all(cat_link, fixed('.'), '-'),
    category = ifelse(tag %in% tags,
      text_spec(compact_category, link=cat_link, color=LINK_COLOR, bold = T),
      compact_category),
    n = paste0(n, ' (', n_secondary, ')'),
    n = str_remove(n, fixed('(NA)'))) 

df_summ_formatted %>%
  select(subject, category, n) %>%
  kable(caption = 'Number of articles by subject and primary category. Colored titles represent hyperlinks that take you below to abstracts. Key - Subject: Computer Science (5) means there were 5 articles with primary tag CS. Category: Machine Learning (cs.LG) N = 8 (16) means there were 8 primary articles with the (cs.LG) tag but 16 articles had it as a secondary tag, so there should be 24 in total. Click this link to be taken to all 24. Only select categories are highlighted because they are of particular interest to applied data scientists.',
        col.names = c('Subject', 'Category', 'N'),
        escape=FALSE) %>%
  kable_styling('striped') %>%
  column_spec(1, bold=T) %>%
  collapse_rows(columns=1, valign='top') 
```

## Articles for Statitstics, Machine Learning Econonmetrics, and Finance
This section contains all articles with any tag of `stat.AP`, `stat.co`, `stat.ML`, `cs.LG`, `q-fin.ST`, `q-fin.EC`, or `econ-EM`. Only the first two sentences are shown - click the links for more detail.

```{r}

```


```{r, results='asis', layout='l-screen-inset'}
df_relevant <- df %>% 
  filter(str_detect(tags, paste0(tags, collapse='|'))) 
# Loop through tags and produce a section for each tag

df_headings <- df_summ_formatted %>%
      select(tag, compact_category, cat_link) %>%
  rename(primary_category = tag)

for (tg in tags){
  df_filtered <- df_relevant %>%
    filter(str_detect(tags, tg))
  if (nrow(df_filtered) > 0){
    # Create markdown anchor: https://talk.commonmark.org/t/anchors-in-markdown/247
    # title will be pulled from details
    row <- df_headings %>%
      filter(primary_category == tg)
    cat(paste('\n\n###', row$compact_category, '{', row$cat_link,' } \n\n'))
    
    df_filtered %>%
      prep_abstracts() %>%
      mutate(primary_category = tg) %>%
      left_join(df_headings, by='primary_category') %>%
      mutate(primary_category = compact_category) %>% 
      select(-compact_category, -cat_link) %>%
      create_html_table() %>%
      print()
  }
}

  
```



## Data Science arXiv by Primary Tag

The tables below show abstracts organized by category with hyperlinks back to the arXiv site. 

```{r, layout='l-page', results='asis'}
# https://rstudio.github.io/DT/#escaping-table-content

subjects <- df_summary$subject %>% unique()
for (sub in subjects){
  cat(paste('\n\n###', sub, '\n\n'))
  # Loop through multiple datatables: https://stackoverflow.com/a/39734221/2138773
  df %>% 
    filter(subject == sub) %>% 
    prep_abstracts() %>%
    create_html_table() %>%
    print()
}
```
